---
title: "Aides Tests"
author: "Damaris Avila"
date: "December 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tigerstats)
require(dplyr)

Scarolina<-read.csv(file="south_carolina_aides.csv",header=TRUE)
ScarolinaF<-filter(Scarolina, sex=="F")
ScarolinaM<-filter(Scarolina, sex=="M")
ScarolinaFtable<-table(ScarolinaF$infected, ScarolinaF$group)
ScarolinaMtable<-table(ScarolinaM$infected, ScarolinaM$group)
ScarolinaALLtable<-table(Scarolina$infected, Scarolina$group)

korea<-read.csv(file="korea_aides.csv",header=TRUE)
koreaF<-filter(korea, sex=="F")
koreaM<-filter(korea, sex=="M")
koreaMtable<-table(koreaM$infected, koreaM$group)
koreaFtable<-table(koreaF$infected, koreaF$group)
koreaALLtable<-table(korea$infected, korea$group)


cenAfrica<-read.csv(file="central_africa_aides.csv",header=TRUE)
cenAfricaF<-filter(cenAfrica, sex=="F")
cenAfricaM<-filter(cenAfrica, sex=="M")
cenAfricaFtable<-table(cenAfricaF$infected, cenAfricaF$group)
cenAfricaMtable<-table(cenAfricaM$infected, cenAfricaM$group)
cenAfricaALLtable<-table(cenAfrica$infected, cenAfrica$group)
```

#Instructions

Three new preventative aides drugs were tested in clinical trials in 2015. The trial designs were modeled after the famous RV_144 Aides Trials in Tailand in 2009.  Analyze and report on the apparent effectiveness of the three different drugs. Does sex of the participant seem to be a factor for any of the drugs?


# South Carolina

We will subset the data by sex to see if the drug is sex specific. Then we will evaluate the group as a whole.    
# #  South Carolina Females  

The following is a bar chart of South Carolina females who were tested in the clinical trials. The graph shows 60% of those females who were infected with AIDS were taking the placebo, while 40% of those infected were taking the tested drug. Of those females who were not infected with AIDS slightly more were taking the drug than those who were taking the placebo.  
```{r}
barchartGC(ScarolinaFtable, type="percent", xlab="Group", ylab="Percent")
```

The following are the row percents of the South Carolina females who were tested in the clinical trials. Of those who were infected 38.83% were taking the test drug and 61.17% were taking the placebo. There is a significant difference between these results, demonstrating the drug's efficacy. Furhtermore, of those who were not infected with AIDS, 50.55% were taking the tested drug, and 49.45% were taking the placebo.
```{r}
rowPerc(ScarolinaFtable)
```

The following is the chi squared test of the table of females who were tested in the clinical trials in South Carolina. 
```{r}
chisqtestGC(ScarolinaFtable, graph= TRUE)
```
*My null hypothesis is:*  
**H~0~** : The vaccine, or drug,  is NOT effective in preventing HIV.  
*My alternate hypothesis is:*  
**H~a~** : The vaccine, or drug, IS effective in preventing HIV.  

The chi squared statistic is 4.9298 which is a measure of how different our two way table is from our expected two way table if our null hypothesis, shown above, was true. This also resulted in a p-value of 0.0264, or a 2.64% chance of getting a chi squared value of 4.9298 or higher if the null was actually true and the drug, vaccine, is NOT effective in preventing AIDS. Because our p-value is lower than 5%, we can reject the null, and say we have shown enough data to show the vaccine is effective to some degree in preventing AIDS in females in South Carolina. 


The following is a bar chart of South Carolina males who were tested in the clinical trials. This bar chart shows about 56% of those infected with AIDS had taken the drug, while just over 40% were taking the placebo. These numbers show a lack of efficacy in the prevention of AIDS. Also, of those who were not infected, percentages were very close to even, about 51% were taking the placebo and 49% were taking the drug. 
```{r}
barchartGC(ScarolinaMtable, type="percent", xlab="Group", ylab="Percent")
```


The following are the row percents of the South Carolina males who were tested in the clinical trials. Supporting the bar chart above, of those infected with AIDS, 57.35% were taking the drug, while 42.65% were taking the placebo. Further supporting the ineffectiveness of the drug. Also, of those not infected 49.52% were taking the drug, while 50.48% were taking the placebo. 
```{r}
rowPerc(ScarolinaMtable)
```

The following is a chi-squared test for the South Carolina males who were tested in the clinical trials. 
```{r}
chisqtestGC(ScarolinaMtable, graph= TRUE)
```

*My null hypothesis is:*  
**H~0~** : The vaccine, or drug,  is NOT effective in preventing HIV.  
*My alternate hypothesis is:*  
**H~a~** : The vaccine, or drug, IS effective in preventing HIV.  

The chi-square test above resulted in a chi-squared statistic of 2.8293 which is a measure of how different our two way
table is from our expected two way table if our null hypothesis, shown above, was true. This test also resulted in a
p-value of 0.0926, or a 9.26% chance of getting a chi-squared value of 2.8293 or higher if the null was actually true and
the drug, vaccine, is NOT effective in preventing AIDS. Since we expected the percentage of people who were infected and
taking the drug, or vaccine, to be less than those taking the placebo, the p-value is high and we are not to reject the
null and have failed to prove the alternate. We can say there has not been enough evidence to disprove the null hypothesis
nor prove the alternate for the South Carolina males who were tested in the clinical trials. 



The following is a bar chart of all the South Carolina individuals who were tested in the clinical trials. In this graph, all the individuals are accounted for for an overall analysis of the drug tested in South Carolina. This graph shows a higher percentage of individuals who were infected with AIDS were taking the placebo, about 51% than those who were taking the drug, about 49%
```{r}
barchartGC(ScarolinaALLtable, type="percent")
```

The following are the row percents of all the South Carolina individuals who were tested in the clinical trials. In support to the bar chart above, of those individuals who were infected 49.37% were taking the tested drug and 50.63% were taking the placebo. These numbers show an inefficacy of the tested drugs for the group overall. 
```{r}
rowPerc(ScarolinaALLtable)
```


The following is a chi-squared test of all the South Carolina individuals who were tested in the clincal trials. 
```{r}
chisqtestGC(ScarolinaALLtable, graph= TRUE)
```

*My null hypothesis is:*  
**H~0~** : The vaccine, or drug,  is NOT effective in preventing HIV.  
*My alternate hypothesis is:*  
**H~a~** : The vaccine, or drug, IS effective in preventing HIV.  

The chi-square test above resulted in a chi-squared statistic of 0.0177 which is a measure of how different our two way
table is from our expected two way table if our null hypothesis, shown above, was true. This test also resulted in a
p-value of 0.8942, or a 89.42% chance of getting a chi-squared value of 0.0177 or higher if the null was actually true and
the drug, vaccine, is NOT effective in preventing AIDS. Since we expected the percentage of people who were infected and
taking the drug, or vaccine, to be less than those taking the placebo, the p-value is high and we are not to reject the
null and have failed to prove the alternate. We can say there has not been enough evidence to disprove the null hypothesis
nor prove the alternate for the South Carolina males who were tested in the clinical trials.


By looking at the group as a whole and the extremely high p-value when analyzing all the South Carolina individuals we can
say the vaccine is definitely not effective in preventive the infection. However, when we seperate the indivudals by sex 
we can see the drug was actually effective to a small degree in the females. There is a big change in numbers when
analyzing the males. There might be a biological reason behind it, but in conclusion, this drug was effective in females, 
not males, and could be a drug that would be beneficial to recommend females take. 


# Central Africa

We will subset the data by sex to see if the drug is sex specific. Then we will evaluate the group as a whole

The following is a bar chart of Central Africa females who were tested in the clinical trials. 
```{r}
barchartGC(cenAfricaFtable, type="percent", xlab="Group", ylab="Percent")
```

The following is a barchart of Central Africa males who were tested in the clinical trials. 
```{r}
barchartGC(cenAfricaMtable, type="percent", xlab="Group", ylab="Percent")
```

The following is a barchart of all Central Africa individuals who were tested in the clinical trials. 

```{r}
barchartGC(cenAfricaALLtable, type="percent", xlab="Group", ylab="Percent")
```




# Korea

We will subset the data by sex to see if the drug is sex specific. Then we will evaluate the group as a whole.

The following is a barchart of Korean females who were tested in the clinical trials. 
```{r}
barchartGC(koreaFtable, type="percent", xlab="Group", ylab="Percent")
```

The following is a bar chart of Korean males who were tested in the clinical trials.   
```{r}
barchartGC(koreaMtable, type="percent", xlab="Group", ylab="Percent")
```

The following is a bar chart of all Korean indiviiduals who were tested in the clinical trials.   
```{r}
barchartGC(koreaALLtable, type="percent", xlab="Group", ylab="Percent")
```

