---
title: "Aides Tests"
author: "Damaris Avila"
date: "December 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tigerstats)
require(dplyr)

Scarolina<-read.csv(file="south_carolina_aides.csv",header=TRUE)
ScarolinaF<-filter(Scarolina, sex=="F")
ScarolinaM<-filter(Scarolina, sex=="M")
ScarolinaFtable<-table(ScarolinaF$infected, ScarolinaF$group)
ScarolinaMtable<-table(ScarolinaM$infected, ScarolinaM$group)
ScarolinaALLtable<-table(Scarolina$infected, Scarolina$group)

korea<-read.csv(file="korea_aides.csv",header=TRUE)
koreaF<-filter(korea, sex=="F")
koreaM<-filter(korea, sex=="M")
koreaMtable<-table(koreaM$infected, koreaM$group)
koreaFtable<-table(koreaF$infected, koreaF$group)
koreaALLtable<-table(korea$infected, korea$group)


cenAfrica<-read.csv(file="central_africa_aides.csv",header=TRUE)
cenAfricaF<-filter(cenAfrica, sex=="F")
cenAfricaM<-filter(cenAfrica, sex=="M")
cenAfricaFtable<-table(cenAfricaF$infected, cenAfricaF$group)
cenAfricaMtable<-table(cenAfricaM$infected, cenAfricaM$group)
cenAfricaALLtable<-table(cenAfrica$infected, cenAfrica$group)
```

#Instructions

Three new preventative aides drugs were tested in clinical trials in 2015. The trial designs were modeled after the famous RV_144 Aides Trials in Tailand in 2009.  Analyze and report on the apparent effectiveness of the three different drugs. Does sex of the participant seem to be a factor for any of the drugs?


# South Carolina

We will subset the data by sex to see if the drug is sex specific. Then we will evaluate the group as a whole.    
# #  South Carolina Females  

The following is a bar chart of South Carolina females who were tested in the clinical trials. The graph shows 60% of those females who were infected with AIDS were taking the placebo, while 40% of those infected were taking the tested drug. Of those females who were not infected with AIDS slightly more were taking the drug than those who were taking the placebo.  
```{r}
barchartGC(ScarolinaFtable, type="percent", xlab="Group", ylab="Percent")
```

The following are the row percents of the South Carolina females who were tested in the clinical trials. Of those who were infected 38.83% were taking the test drug and 61.17% were taking the placebo. There is a significant difference between these results, demonstrating the drug's efficacy. Furhtermore, of those who were not infected with AIDS, 50.55% were taking the tested drug, and 49.45% were taking the placebo. 
```{r}
rowPerc(ScarolinaFtable)
```

The following is the chi squared test of the table of females who were tested in the clinical trials in South Carolina. 
```{r}
chisqtestGC(ScarolinaFtable, graph= TRUE)
```
*My null hypothesis is:*  
**H~0~** : The vaccine is NOT effective in preventing HIV.  
*My alternate hypothesis is:*  
**H~a~** : The vaccine IS effective in preventing HIV.  

The chi squared statistic is 4.9298 which is a measure of how different our two way table is from our expected two way table if our null hypothesis, shown above, was true. This also resulted in a p-value of 0.0264, or a 2.64% chance of getting a chi squared value of 4.9298 or higher if the null was actually true and the drug, vaccine, is NOT effective in preventing AIDS. Because our p-value is lower than 5%, we can reject the null, and say we have shown enough data to show the vaccine is effective to some degree in preventing AIDS. 


The following is a bar chart of South Carolina males who were tested in the clinical trials. 
```{r}
barchartGC(ScarolinaMtable, type="percent", xlab="Group", ylab="Percent")
```


The following is a bar chart of all the South Carolina individuals who were tested in the clinical trials.
```{r}
barchartGC(ScarolinaALLtable, type="percent")
```






# Central Africa

We will subset the data by sex to see if the drug is sex specific. Then we will evaluate the group as a whole

The following is a bar chart of Central Africa females who were tested in the clinical trials. 
```{r}
barchartGC(cenAfricaFtable, type="percent", xlab="Group", ylab="Percent")
```

The following is a barchart of Central Africa males who were tested in the clinical trials. 
```{r}
barchartGC(cenAfricaMtable, type="percent", xlab="Group", ylab="Percent")
```

The following is a barchart of all Central Africa individuals who were tested in the clinical trials. 

```{r}
barchartGC(cenAfricaALLtable, type="percent", xlab="Group", ylab="Percent")
```




# Korea

We will subset the data by sex to see if the drug is sex specific. Then we will evaluate the group as a whole.

The following is a barchart of Korean females who were tested in the clinical trials. 
```{r}
barchartGC(koreaFtable, type="percent", xlab="Group", ylab="Percent")
```

The following is a bar chart of Korean males who were tested in the clinical trials.   
```{r}
barchartGC(koreaMtable, type="percent", xlab="Group", ylab="Percent")
```

The following is a bar chart of all Korean indiviiduals who were tested in the clinical trials.   
```{r}
barchartGC(koreaALLtable, type="percent", xlab="Group", ylab="Percent")
```

